---
title: "R Notebook"
output: html_notebook
---

#load Tidyverse library and download data from Github TidyTuesday repository
```{r}
library(tidyverse)
tv_shows<-readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-08/IMDb_Economist_tv_ratings.csv")
```
#rename the "title" variable to "Program"
```{r}
colnames(tv_shows)[colnames(tv_shows)=="title"] <- "Program"
```
#sort data by season number in descending order
```{r}
tv_seasons<-tv_shows%>%
     arrange(desc(seasonNumber))
```
#filter tv shows with more than 10 seasons and group by Program
```{r}
tv_longrun<-tv_seasons%>%
     filter(seasonNumber>=10)%>%
    group_by(Program)
```
#identify unique Programs by titleID
```{r}
tv_longrun_unique<-distinct(tv_longrun,titleId)
```
#filter all seasons of Programs that lasted at least 10 seasons
```{r}
tv_lr<-tv_shows%>%
     filter(Program %in% tv_longrun_unique$Program)
```
#plot to compare rating by season for Programs that lasted at least 10 seasons
```{r}
ggplot()+
     geom_line(data=tv_lr,mapping=aes(x=seasonNumber,y=av_rating,color=Program))+
      theme_bw()+
      xlab("Season")+
      ylab("Average Rating")
```
#filter out long running comedy series
```{r}
comedy<-tv_lr%>%
     filter(str_detect(tv_lr$genres,"Comedy"))
```
#plot to compare rating by season among comedy Programs
```{r}
ggplot()+
     geom_line(data=comedy,mapping=aes(x=seasonNumber,y=av_rating,color=Program),na.rm=TRUE)+
      theme_bw()+
      xlim(0,15)+
      ylim(5,9)+
      xlab("Season")+
      ylab("Average Rating")
```
#split genres variable so Programs with more than one genre appear once for each genre
```{r}
genre<-separate_rows(tv_shows,genres,sep=",")%>%
  arrange(genres)%>%
  group_by(genres, seasonNumber)%>%
  summarize(mean=mean(av_rating))
```
#plot to compare rating per season among all genres
```{r}
ggplot()+
  geom_line(data=genre,mapping=aes(x=seasonNumber,y=mean,color=genres))+
  theme_bw()+
  labs(title="Average Rating per Season by Genre",x="Season",y="Average Rating")
```
#filter Programs that lasted less than 10 seasons and are not Reality TV Programs
```{r}
genre_short<-genre%>%
  filter(seasonNumber<=10)%>%
  filter(!str_detect(genres,"Reality"))
```
#plot to compare rating per season among short lived, non Reality TV Programs
```{r}
ggplot()+
  geom_line(data=genre_short,mapping=aes(x=seasonNumber,y=mean,color=genres))+
  theme_bw()+
  labs(title="Average Rating per Season (<10) by Genre",x="Season",y="Average Rating")
```

